---
title: "Explore Mercury Data"
author: "Dr. Gordon McDonald"
date: "19/08/2020"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache = T)
library(tidyverse)
library(data.table)
library(plotly)
library(skimr)
library(GGally)
#load mercury_data
load(file = here::here("Data/mercury_data_exported.Rdata"))
```

# Skim mercury data

Have a look at each of the variables

```{r}
skim(mercury_data)
```

# How many of each kind of fish do we have?{.tabset}

## Table
```{r}
(fish_summary <-
    mercury_data %>% 
  mutate(`Number of Samples` = if_else(is.na(`Number of Samples`),1,`Number of Samples`)) %>% 
  group_by(`Type of Fish`) %>% 
  summarise(Samples = sum(`Number of Samples`),
            mean_conc = round( mean(`Conc Hg Fish [ug/g]`),3),
            sd_conc =  round( sd(`Conc Hg Fish [ug/g]`),3),
            .groups = "drop") %>% 
   {mutate(.,`Type of Fish` = factor(x = `Type of Fish`, 
                                     levels = `Type of Fish`[order(Samples,decreasing = T)], 
                                     ordered = T))} %>% 
   arrange(desc(Samples))) %>% 
  DT::datatable()
```

## Bar graph (All)

```{r}
fish_summary %>% 
  ggplot(aes(x = `Type of Fish`,y = Samples))+
  geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90,hjust = 1,size = 4))
```

## Bar graph (Top 15)

```{r}
fish_summary %>% 
  filter(as.numeric(`Type of Fish`)<=15) %>% 
  {ggplot(.,aes(x = `Type of Fish`,y = Samples))+
  geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90,hjust = 1, size = 7))} %>% ggplotly(height = 500)
```


```{r, message = F, warning = F, eval = F, include = F}
ggpairs(mercury_data %>% 
          select(where(is.numeric)) %>% 
          select(-c("Published", 
                    "Data collected (first)",
                    "Data collected (last)",
                    "Latitude","Longitude"))
)
```

# Fish Size/Age/Weight{.tabset}

## Fish Length{.tabset}
### All

```{r, message = F, warning = F}
mercury_data %>% 
  {ggplot(.,aes(x = `Length of Fish (cm)`))+
      geom_histogram()} %>% 
  ggplotly
```

### Small Fish Length
Length < 1m
```{r, message = F, warning = F}
mercury_data %>% 
  filter(`Length of Fish (cm)`<100) %>% 
  {ggplot(.,aes(x = `Length of Fish (cm)`))+
      geom_histogram()} %>% 
  ggplotly
```

## Fish Weight{.tabset}
### All

```{r, warning = F, message=F}
mercury_data %>% 
  {ggplot(.,aes(x = `Weight of Fish (g)`))+
      geom_histogram()} %>% 
  ggplotly
```

### Small Fish Weight
Weight < 10kg
```{r, warning = F, message=F}
mercury_data %>% 
  filter(`Weight of Fish (g)`< 10000) %>% 
  {ggplot(.,aes(x = `Weight of Fish (g)`))+
      geom_histogram()} %>% 
  ggplotly
```

## Fish Age{.tabset}

### All

```{r, warning = F, message = F}
mercury_data %>% 
  {ggplot(.,aes(x = `Age (years)`))+
      geom_histogram()} %>% 
  ggplotly
```

### Young Fish Age
Age < 10 years
```{r, warning = F, message = F}
mercury_data %>% 
  filter(`Age (years)`<10) %>% 
  {ggplot(.,aes(x = `Age (years)`))+
      geom_histogram()} %>% 
  ggplotly
```

## Fish Size/Weight{.tabset}

### All

```{r, message = F, warning = F}
mercury_data %>%
  select(`Length of Fish (cm)`,
         `Weight of Fish (g)`,
         `Type of Fish`) %>%
  na.omit %>%
  {ggplot(.,aes(x = `Length of Fish (cm)`,
                y = `Weight of Fish (g)`,
                color = `Type of Fish`))+
      geom_point()+
      coord_cartesian(xlim = c(0,300))+
      theme(legend.position = 'none')} %>% 
  ggplotly
```

### Big Fish size/weight (Tuna/Shark)

```{r, message = F, warning = F}
mercury_data %>%
  select(`Length of Fish (cm)`,
         `Weight of Fish (g)`,
         `Type of Fish`) %>%
  na.omit %>%
  filter(`Length of Fish (cm)`>100,
         `Weight of Fish (g)`>10000) %>% 
  # filter(grepl("(tuna|shark)",
  #              as.character(`Type of Fish`),
  #              ignore.case = TRUE)) %>% 
  {ggplot(.,aes(x = `Length of Fish (cm)`,
                y = `Weight of Fish (g)`,
                color = `Type of Fish`))+
      geom_point()+
      coord_cartesian(xlim = c(0,300))+
      theme(legend.position = 'none')} %>% 
  ggplotly
```

### Small Fish size/weight

```{r, message = F, warning = F}
mercury_data %>%
  select(`Length of Fish (cm)`,
         `Weight of Fish (g)`,
         `Type of Fish`) %>%
  na.omit %>%
  filter(`Length of Fish (cm)`<=100,
         `Weight of Fish (g)`<=10000) %>% 
  # filter(!grepl("(tuna|shark)",
  #               as.character(`Type of Fish`),
  #               ignore.case = TRUE)) %>% 
  {ggplot(.,aes(x = `Length of Fish (cm)`,
                y = `Weight of Fish (g)`,
                color = `Type of Fish`))+
      geom_point()+
      theme(legend.position = 'none')} %>% 
  ggplotly
```

## Fish age/weight

```{r, message = F, warning = F}
mercury_data %>%
  select(`Age (years)`,
         `Weight of Fish (g)`,
         `Type of Fish`) %>%
  na.omit %>% 
  {ggplot(.,aes(x = `Age (years)`,
                y = `Weight of Fish (g)`,
                color = `Type of Fish`))+
      geom_point()+
      theme(legend.position = 'none')} %>% 
  ggplotly
```

## Fish age/length

```{r, message = F, warning = F}
mercury_data %>%
  select(`Age (years)`,
         `Length of Fish (cm)`,
         `Type of Fish`) %>%
  na.omit %>% 
  {ggplot(.,aes(x = `Age (years)`,
                y = `Length of Fish (cm)`,
                color = `Type of Fish`))+
      geom_point()+
      theme(legend.position = 'none')} %>% 
  ggplotly
```

# Concentration of Mercury{.tabset}

## Fish Conc{.tabset}

### All

```{r, message = F, warning = F}
mercury_data %>% 
  {ggplot(.,aes(x = `Conc Hg Fish [ug/g]`))+
      geom_vline(xintercept = 0.5, color = "red")+
      geom_histogram()} %>% 
  ggplotly
```

### Excluding highest

Everything below 4 ug/g

```{r, message = F, warning = F}
mercury_data %>% 
  filter(`Conc Hg Fish [ug/g]`<4) %>% 
  {ggplot(.,aes(x = `Conc Hg Fish [ug/g]`))+
      geom_vline(xintercept = 0.5, color = "red")+
      geom_histogram()} %>% 
  ggplotly
```

## Sediment Conc{.tabset}

### All
```{r, message = F, warning = F}
mercury_data %>% 
  {ggplot(.,aes(x = `Conc Hg Sediment [ug/g]`))+
      geom_histogram()} %>% 
  ggplotly
```

### Excluding highest

Everything below 1 ug/g

```{r, message = F, warning = F}
mercury_data %>% 
  filter(`Conc Hg Sediment [ug/g]`<1) %>% 
  {ggplot(.,aes(x = `Conc Hg Sediment [ug/g]`))+
      geom_histogram()} %>% 
  ggplotly
```

## Fish conc/length

```{r, message = F, warning = F}
mercury_data %>%
  select(`Conc Hg Fish [ug/g]`,
         `Length of Fish (cm)`,
         `Type of Fish`) %>%
  na.omit %>% 
  {ggplot(.,aes(x = `Conc Hg Fish [ug/g]`,
                y = `Length of Fish (cm)`,
                color = `Type of Fish`))+
      geom_vline(xintercept = 0.5,color = "red")+
      geom_point()+
      theme(legend.position = 'none')} %>% 
  ggplotly
```




## Fish conc/weight

```{r, message = F, warning = F}
mercury_data %>%
  select(`Conc Hg Fish [ug/g]`,
         `Weight of Fish (g)`,
         `Type of Fish`) %>%
  na.omit %>% 
  {ggplot(.,aes(x = `Conc Hg Fish [ug/g]`,
                y = `Weight of Fish (g)`,
                color = `Type of Fish`))+
      geom_vline(xintercept = 0.5,color = "red")+
      geom_point()+
      theme(legend.position = 'none')} %>% 
  ggplotly
```

## Fish conc/Age

```{r, message = F, warning = F}
mercury_data %>%
  select(`Conc Hg Fish [ug/g]`,
         `Age (years)`,
         `Type of Fish`) %>%
  na.omit %>% 
  {ggplot(.,aes(x = `Conc Hg Fish [ug/g]`,
                y = `Age (years)`,
                color = `Type of Fish`))+
      geom_vline(xintercept = 0.5,color = "red")+
      geom_point()+
      theme(legend.position = 'none')} %>% 
  ggplotly
```


## Fish conc/Year of Publication

```{r, message = F, warning = F}
mercury_data %>%
  select(`Conc Hg Fish [ug/g]`,
         Published,
         `Type of Fish`) %>%
  na.omit %>% 
  {ggplot(.,aes(x = `Conc Hg Fish [ug/g]`,
                y = Published,
                color = `Type of Fish`))+
      geom_vline(xintercept = 0.5,color = "red")+
      geom_point()+
      theme(legend.position = 'none')} %>% 
  ggplotly
```



## Fish conc/Sediment conc

```{r, message = F, warning = F}
mercury_data %>%
  select(`Conc Hg Fish [ug/g]`,
         `Conc Hg Sediment [ug/g]`,
         `Type of Fish`) %>%
  na.omit %>% 
  {ggplot(.,aes(x = `Conc Hg Fish [ug/g]`,
                y = `Conc Hg Sediment [ug/g]`,
                color = `Type of Fish`))+
      geom_vline(xintercept = 0.5,color = "red")+
      geom_point()+
      theme(legend.position = 'none')+
      coord_cartesian(expand = FALSE)} %>% 
  ggplotly
```



## Fish conc/Region

```{r, message = F, warning = F}
mercury_data %>%
  select(`Conc Hg Fish [ug/g]`,
         Region) %>%
  na.omit %>% 
  {ggplot(.,aes(x = Region,
                y = `Conc Hg Fish [ug/g]`))+
      geom_hline(yintercept = 0.5,color = "red")+
      geom_boxplot()+
      theme(legend.position = 'none')+
      coord_cartesian(expand = FALSE)}
```



## Fish conc/Type of Fish{.tabset}

### All

```{r, message = F, warning = F, fig.height = 7}
mercury_data %>%
  mutate(`Type of Fish` = factor(x = `Type of Fish`,levels = fish_summary$`Type of Fish`[order(fish_summary$mean_conc,decreasing = T)])) %>% 
  select(`Conc Hg Fish [ug/g]`,
         `Type of Fish`) %>%
  na.omit %>% 
  {ggplot(.,aes(x = `Type of Fish`,
                y = `Conc Hg Fish [ug/g]`,
                text = `Type of Fish`))+
      geom_hline(yintercept = 0.5,color = "red")+
      geom_boxplot()+
      theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6),
            legend.position = 'none')+
      coord_cartesian(expand = FALSE)} %>% 
  ggplotly(tooltip = "text")
```

### Top 20

```{r, message = F, warning = F, fig.height = 7}
mercury_data %>%
  mutate(`Type of Fish` = factor(x = `Type of Fish`,
                                 levels = fish_summary$`Type of Fish`[order(fish_summary$mean_conc,
                                                                            decreasing = T)])) %>% 
  select(`Conc Hg Fish [ug/g]`,
         `Type of Fish`) %>%
  na.omit %>% 
  filter(as.numeric(`Type of Fish`)<=20) %>% 
  {ggplot(.,aes(x = `Type of Fish`,
                y = `Conc Hg Fish [ug/g]`,
                text = `Type of Fish`))+
      geom_hline(yintercept = 0.5,color = "red")+
      geom_boxplot()+
      theme(axis.text.x = element_text(angle = 90, hjust = 1),
            legend.position = 'none')+
      coord_cartesian(expand = FALSE)} %>% 
  ggplotly(tooltip = "text")
```